SPARQL on the Open, Decentralised Web

### 2016-07-22 dissertation/related.tex
Helen's corrections

There is a substantial amount of relevant literature and ongoing
research that is well aligned with the present work, either because
the authors share parts of the motivation, or more serendipitous
reuse.

The related work on the topic of evaluations based on contemporary
statistics, and on the topic of philosophy is scarce, and have been
dealt with in the papers. For the other topics dealt with in this
thesis, a more extensive treatment is needed to place this work in
context:

@@

An example of this use is the work considered in
Section~\ref{sec:conlapis}, another example is the new direction
proposed by Verborgh et al. \cite{ldf1} who wanted a system that would
be able to answer individual triple patterns to enable clients to
answer queries, and to do so, they employed hypermedia, and formulated
this as Triple Pattern Fragments. The key to this hypermedia is that
an answer does not only include the data, but also metadata,
importantly a cardinality estimate for the triple pattern, and also
control information, that tells a client explicitly in every response
how further interaction should be done, like obtaining further data.

@@

However, it is not necessarily the case that evaluating the entire
query on the client side is the most efficient, depending in
particular on the perspective of the different
actors. Chapter~\ref{sec:tpfcacheplanning} adopts Triple Pattern
Fragments, but makes different assumptions, e.g. tries to avoid a
large number of HTTP requests.

Also, TPF mandates that every response must contain a cardinality
estimate for every triple pattern, an operation that may be quite
expensive for the server, and
%TODO: ref characteristic sets?
indeed, this information may be so expensive to provide that SPARQL
planners may not provide it to keep the cost of planning itself
down.

@@

HERMES \cite{adali1996query} provided motivation for the paper in
Section~\ref{sec:conpush}, but is also important prior art for the
work in Chapter~\ref{sec:tpfcacheplanning}. An important concept in
the HERMES system is ``invariants''. They are, in our terminology,
patterns that do not need to be looked up remotely, but can be
integrated in the query answer by the mediator. The expectation in
HERMES is that such invariants will be encoded by a domain expert. The
Attean framework, which is used extensively in this work and is
detailed in Section~\ref{sec:attean} provides facilities to do this in
an easy and extensible way, and also accommodates for the task of
computing a scalar cost metric, but HERMES has a more advanced cost
model in that it has a cost vector with estimates for time required to
find the first answer, time to find all answers and the cardinality of
the answer set. Caching summary statistics is another important
feature of HERMES, and it also contributes lossy and lossless
summaries, for the purpose of saving space and reduce the time it
takes to compute a cost estimate. HERMES further has a query rewriter
and optimiser, that rewrites the query taking into account the cache
and the invariants. While HERMES is more general than what I detail in
Chapter~\ref{sec:tpfcacheplanning}, we note that while invariants
could be coded in Attean, a SPARQL query cache should be oblivious to
domain knowledge and therefore shouldn't assume any invariants, except
in the possible but unlikely case where the remote server declares
that an infinite freshness lifetime for the result of a certain graph
pattern. While statistics summarisation is relevant and should be
studied in future work, we note that with a Linked Data Fragments
server, a cardinality estimate is required to be available for every
triple pattern, and is also cacheable with the same constraints as the
data themselves.

Interestingly, \cite{papailiou2015graph} opens by noting that many
optimisations used by SQL databases are ineffective in RDF databases,
due to that RDF schema, even if they exist, are not as helpful as the
obligatory schema in SQL databases. This provides an opportunity for
novel research. The solution proposed in this paper is not a cache on
a proxy or a mediator like my primary focus is. The execution engine
in this case, has access to both the primary indexes of the RDF store
and the cache. A key contribution of the paper is a canonical
labelling algorithm that can create a string that will be unique for a
Basic Graph Pattern including optional blocks. This is done by
transforming a SPARQL query to a directed vertex-coloured graph, where
each triple pattern is represented by a vertex, and triple patterns
that share a common variable is linked with an edge. For each vertex,
a label that consists of three IDs are created. Bound terms are
translated to IDs using a dictionary, while variables are translated
to zero. Then, the graph is directed and edges labelled according to
the type of join they represent. The vertex and edge labels are then
translated to non-negative integers (colours) using a sort key, to
form an directed vertex-and-edge-coloured graph that is the authors
claim is isomorphic with the original SPARQL. Finally, the graph is
transformed to a simpler directed vertex-coloured graph. With this
graph, the author's method can produce a canonical label for the query
graph.

@@

Acosta et al. has published a poster on ``SHEPHERD''
\cite{acosta2014shepherd}, presumed to be ongoing work, to
create a SPARQL processor that takes into account observations done by
the SPARQLES survey \cite{buil2013sparql}, to decompose a SPARQL query into
subqueries that is indicated to be easier for the server to evaluate,
and so is well aligned with our goal to ease the load of the
server. It is not quite clear from the brief description in 
\cite{acosta2014shepherd} whether the SHEPHERD system can cache results locally.

@@

Naturally, the problem of caching in a proxy has much in common with
the federation problem as the cache, as seen from the query execution
engine's perspective, could be viewed as yet another possible data
source, and the number of HTTP requests to the remote endpoint can be
wisely assumed to be kept small. Schwarte et
al. \cite{springerlink:10.1007/978-3-642-25073-6-38} addressed these
problems with FedX, in which they sought to group connected subgraphs
with their concept of exclusive groups.  Montoya et
al. \cite{montoya2015federated} propose a query federation system that
minimise the data needed to be transferred while preserving
completeness in the common case where the data relevant to a query is
available on different endpoints. Görlitz and Staab \cite{splendid}
exploited cardinality estimates sometimes available in VoID
\cite{voidnote} descriptions to improve join order execution plans. It
could be advantageous for the system in
Chapter~\ref{sec:tpfcacheplanning} to exploit as well, and
straightforward to implement, but since it has not seen wide uptake
and the statistics is mandated by Triple Pattern Fragments, and
therefore available in some cases to the planner, we chose not to.

>>>>

There is a substantial amount of relevant literature and ongoing
research that are well aligned with the present work, either because
the authors share parts of the motivation, or this work finds serendipitous
reuse of prior work.

The related work on the topic of evaluations based on contemporary
statistics, and on the topic of philosophy is scarce, and has been
dealt with in the papers. For the other topics dealt with in this
thesis, a more extensive treatment is needed to place this work in
context:

@@

An example of this use is the work considered in
Section~\ref{sec:conlapis}, another example is the new direction
proposed by Verborgh et al. \cite{ldf1} who wanted a system that would
be able to answer individual triple patterns to enable clients to
answer queries, and in doing so, they employed hypermedia, and formulated
this as Triple Pattern Fragments. The key to this hypermedia is that
an answer does not only include the data, but also metadata,
importantly a cardinality estimate for the triple pattern, and also
control information, that tells a client explicitly in every response
how further interaction should be done, like in obtaining further data.

@@

However, it is not necessarily the case that evaluating the entire
query on the client side is the most efficient, depending in
particular on the perspective of the different
actors. Chapter~\ref{sec:tpfcacheplanning} adopts Triple Pattern
Fragments, but makes different assumptions, for example by trying to avoid a
large number of HTTP requests.

Also, TPF mandates that every response must contain a cardinality
estimate for every triple pattern, an operation that may be quite
expensive for the server.
%TODO: ref characteristic sets?
Indeed, this information may be so expensive to provide that SPARQL
planners may not do so, in order to keep the cost of planning itself
down.

@@

HERMES \cite{adali1996query} provided motivation for the paper in
Section~\ref{sec:conpush}, but is also important for the
work in Chapter~\ref{sec:tpfcacheplanning}. An important concept in
the HERMES system is ``invariants''. They are, in our terminology,
patterns that do not need to be looked up remotely, but can be
integrated in the query answer by the mediator. The expectation in
HERMES is that such invariants will be encoded by a domain expert. The
Attean framework, which is used extensively in this work and is
detailed in Section~\ref{sec:attean} provides facilities to do this in
an easy and extensible way, and also accommodates for the task of
computing a scalar cost metric, but HERMES has a more advanced cost
model in that it has a cost vector with estimates for time required to
find the first answer, time to find all answers and the cardinality of
the answer set. Caching summary statistics is another important
feature of HERMES, and it also contributes lossy and lossless
summaries, for the purpose of saving space and reduce the time it
takes to compute a cost estimate. HERMES further has a query rewriter
and optimiser, that rewrites the query taking into account the cache
and the invariants. While HERMES is more general than what I detail in
Chapter~\ref{sec:tpfcacheplanning}, we note that while invariants
could be coded in Attean, a SPARQL query cache should be oblivious to
domain knowledge and therefore shouldn't assume any invariants, except
in the possible but unlikely case where the remote server declares
that an infinite freshness lifetime for the result of a certain graph
pattern. While statistics summarisation is relevant and should be
studied in future work, we note that with a Linked Data Fragments
server, a cardinality estimate is required to be available for every
triple pattern, and is also cacheable with the same constraints as the
data themselves.

Interestingly, \cite{papailiou2015graph} opens by noting that many
optimisations used by SQL databases are ineffective in RDF databases,
due to that RDF schema, even if they exist, are not as helpful as the
obligatory schema in SQL databases. This provides an opportunity for
novel research. The solution proposed in this paper is not a cache on
a proxy or a mediator like that which is my primary focus. The execution engine
in this case, has access to both the primary indexes of the RDF store
and the cache. A key contribution of the paper is a canonical
labelling algorithm that can create a string that will be unique for a
Basic Graph Pattern including optional blocks. This is done by
transforming a SPARQL query to a directed vertex-coloured graph, where
each triple pattern is represented by a vertex, and triple patterns
that share a common variable is linked with an edge. For each vertex,
a label that consists of three IDs are created. Bound terms are
translated to IDs using a dictionary, while variables are translated
to zero. Then, the graph is directed and edges labelled according to
the type of join they represent. The vertex and edge labels are then
translated to non-negative integers (colours) using a sort key, to
form a directed vertex-and-edge-coloured graph that the authors
claim is isomorphic with the original SPARQL. Finally, the graph is
transformed to a simpler directed vertex-coloured graph. With this
graph, the author's method can produce a canonical label for the query
graph.

@@

Acosta et al. has published a poster on ``SHEPHERD''
\cite{acosta2014shepherd}, presumed to be ongoing work, to
create a SPARQL processor that takes into account observations done by
the SPARQLES survey \cite{buil2013sparql}, to decompose a SPARQL query into
subqueries that are indicated to be easier for the server to evaluate,
and so is well aligned with our goal to ease the load of the
server. It is not quite clear from the brief description in 
\cite{acosta2014shepherd} whether the SHEPHERD system can cache results locally.

@@

Naturally, the problem of caching in a proxy has much in common with
the federation problem as the cache, as seen from the query execution
engine's perspective, could be viewed as yet another possible data
source, and the number of HTTP requests to the remote endpoint can be
wisely assumed to be kept small. Schwarte et
al. \cite{springerlink:10.1007/978-3-642-25073-6-38} addressed these
problems with FedX, in which they sought to group connected subgraphs
with their concept of exclusive groups.  Montoya et
al. \cite{montoya2015federated} propose a query federation system that
minimise the data needed to be transferred while preserving
completeness in the common case where the data relevant to a query is
available on different endpoints. Görlitz and Staab \cite{splendid}
exploited cardinality estimates sometimes available in VoID
\cite{voidnote} descriptions to improve join order execution plans. It
could be advantageous for the system in
Chapter~\ref{sec:tpfcacheplanning} to exploit as well, and
straightforward to implement, but since it has not seen wide uptake
and the statistics are mandated by Triple Pattern Fragments, and
therefore available in some cases to the planner, we chose not to.
